Predicting Success of Motivational Interviews with Recurrent Neural Network and Probabilistic Model for Patient-Provider Communication Sequences

### 2017-09-18 paper-amia-summit-2018/amia_summit_2018.tex
update reference
update introduction

In general, there are three categories of classification methods for teporally ordered sequences, e.g.: feature-based, distance metric-based and model-based method. Feature-based classification methods transform a sequence into a feature vector and apply a standard supervised learning algorithm, such as support vector machine \cite{leslie2004fast} or decision tree \cite{chuzhanova1998feature}. Shapelet \cite{ye2009time} and pattern \cite{kudenko1998feature, lesh1999mining} based techniques as well as hierarchical approaches \cite{nallam2016effective} have been proposed instead of standard classifiers as well. Distance-based methods measure the similarity between sequences of behavior codes to determine the quality of classification. The most commonly used distance function is Euclidian distance \cite{keogh2003need} with Dynamic Time Wrapping \cite{keogh2000scaling} used for more flexible matching in time series data. The order of observations in a sequence is challenging to model using feature-based methods because the traditional bag-of-words representation of features does not account for the order of observation. At the same time, distance-based methods are more appropriate for time series data. Model-based classification methods train a model by data and then applies it for the classification of new datasets.

@@

One of the model-based classification methods first creates a probabilistic model of a behavioral code sequence, such as Markov Chain (MC) and Hidden Markov Model \cite{rabiner1989tutorial} (HMM). Markov models have limitations when applied to sequence classification because they cannot preserve long-term dependencies in sequences. Similarly, the hidden states of HMMs follow a first-order markov chain and can select only one hidden state at each time step. Therefore, Markov chain and HMMs are only capable of passing upto five preceding steps among sequence steps[ref]\cite{}. On the other hand, Recurrent Neural Networks (RNNs) have no such limitations in preserving long-term dependencies among sequence steps due to their unlimited memory. It was shown that RNNs are very effective in sequence modeling with temporal data. Excellent performance in a wide range of sequence modeling problems with temporal data has been observed using RNNs \cite{nion2013handwritten, lipton2015learning, choi2016doctor}. A variant of RNN called Long Short-Term Memory was proposed \cite{graves2013speech} which solved the vanishing gradient problem \cite{bengio1993problem}.LSTM showed excellent performance in several domain, including speech recognition \cite{graves2013speech}. A further simplification and improvement of the LSTM model, called the Gated Reccurent Unit\cite{chung2014empirical} (GRU), was later proposed.
%Recently, two variant of RNNs is proposed: Long Short-Term Memory\cite{graves2013speech} (LSTM) and an enhanced version of LSTM . 
These variants show markedly improved performance among all RNN variants and as such, this work focuses on these varians of RNN specifically.

In this paper, we test the applicability of state of the art models for the classification of patient-provider communication sequences in the context of clinical interviews. Specifically, we focus on the transcripts of Motivational Interviews (MI) with obese adolescents and their caregivers. Childhood obesity is a serious public health concern in the United States and worldwide. Recent estimates indicate that approximately one third (31.8\%) of US children age 2-19 years are overweight and 16.9\% are obese \cite{ogden2012prevalence}. Adolescents who are obese likely continue to be obese in adulthood and have a greater risk of heart disease, type 2 diabetes, stroke, cancer, and osteoarthritis \cite{general2010surgeon}. One approach to effective obesity intervention is Motivational Interviewing, an evidence-based counseling technique to increase intrinsic motivation and self-efficacy for health-related behavior change. The goal of MI is to encourage patients to explore their own desires, ability, reasons, need for and commitment to the targeted behavior change. These statements, referred to as ``change talk'' (or CT), consistently predict actual behavior change\cite{apodaca2009mechanisms} that can be sustained for as long as 34 months\cite{walker2011influence} after an interview. However, the ability of counselors to consistently elicit this type of patient communication requires knowledge of effective communication strategies for a variety of patients, which can only be obtained through analysis of a large number of annotated interviews. Since manual examination and sequential pattern analysis of transcripts is a very time-consuming process, tailoring of MI interventions to particular populations can take years. Therefore, there is a need for informatics-based methods to facilitate the development of effective MI-based interventions, in general, and theoretically-grounded computational models to explore the mechanisms of MI's efficacy, in particular.

@@

\section*{Methods}
\subsection*{\textit{Data collection}}
The experimental dataset for this work was constructed from the transcripts of 129 motivational interviews, which include a total of 50,239 segmented and annotated utterances. Each transcript consists of a MI interview session involving counselor, adolescent, and caregiver. The utterances were annotated based on MYSCOPE codebook \cite{carcone2013provider}, which are grouped into client codes (adolescent and caregiver) group and counselor codes group. Utterances were divided into successful and unsuccessful communication sequences. Successful communication sequences result in positive change talk and commitment language statements by an adolescent or caregiver, while unsuccessful sequences are the ones that result in negative change talk or commitment language and the sequences, in which no change talk or commitment language statements occur. Out of 5143 observed sequences, 4225 were positive and 918 were negative. Here, the successful sequences had an average length of 9.79 while that of unsuccessful sequences was 9.65. For each of the probabilistic models (Markov chain and HMM), two models were trained, one model was trained using successful sequences and another one was trained using unsuccessful sequences.

@@

Our data set is imbalanced and predominately composed of successful sequences with only a small percentage (17.85\%) of unsuccessful sequences. Usually, predictive accuracy misrepresents the performance of an employed algorithm for an imbalanced data. Because simply predict the majority class (successful) would provide a predictive accuracy 82.15\%. Therefore, it is important to handle imbalance data properly. Many solutions are proposed to deal with the imbalanced data sets, which can be divided into two categories: data level and algorithmic level. In this study, we utilized under and over sampling methods at data processing level for balancing the adolescent obesity data set. Synthetic Minority Over-sampling (SMOTE) method is the most widely used method for over-sampling an imbalanced data set, in which new synthetic examples were generated from minority class \cite{chawla2002smote}. We generated synthetic examples along the borderline between minority examples and their selected nearest neighbors \cite{nguyen2011borderline}. On the other hand, under sampling method under samples the majority class by replacing a cluster of majority samples by the cluster centroid of a K-Means algorithm.  A fragment of an adolescent session transcript is presented in Table~\ref{tab:anno_examp}. Hence, annotation column shows the successful sequence of behavior codes from top to bottom, in particular, ($SS\rightarrow OQO\rightarrow HUPO\rightarrow OQTBN\rightarrow CHT+$) is a successful behavior sequence, where counselor starts with a structured session and gets positive change talk at the end. It was noticed that same context ``Yeah" and ``Yes'' represent different behavior codes CHT+ and HUPW, respectively. Because human brain can remember the past behaviors, and classify the current context into different behavior codes after considering their past behaviors. For this reason, we only consider those models that can capture the past events for the classification of behavior sequences.\\

@@

The above model also refered as first-order MC, since it only considers immediately preceding behavior code when computing the state transition probabilities. In our experiment, we also considered second-order Markov model that compute transition probabilities based on previous two states.

\textbf {Hidden Markov Model (HMM)} is another probabilistic model, which is used for modeling processes varying in time. HMMs are widely used for sequence analysis because of their ability to incorporate dependencies among elements in sequence. Mathmatically, HMM can be defined as the following: $\lambda = (N, M, A, B, \pi)$
\begin{itemize}
\item N is a number of hidden states in the model
\item M is a number of distinct observations symbols per state, i.e. the discrete vocabulary size
\item A is an $N\times N$ state transition probability distribution matrix $A = \{a_{ij}\}$
\item B is an $N\times M$ matrix $B = \{b_j(k)\}$ with observation symbol probability distribution for each state 
\item $\pi$ is the initial state distribution vector $\pi = \{\pi_i\}$
\end{itemize}
HMM is also represented by a compact notation: $\lambda = (A, B, \pi)$ after excluding the structure parameters M and N. The key difference between HMM and MC is that HMM requires specifying the number of hidden states as a model parameter and then the model deduces a sequence of hidden states that best explains the observations along with state transition probabilities and distributions of observation symbols for each state. The Baum-Welch algorithm was used to estimate the parameters of HMMs for successful and unsuccessful interviews using the corresponding training set, while the Viterbi algorithm was used to determine the most likely sequence of hidden states for a given sequence of observations. After assignment of hidden states, the log-likelihood of successful outcome can be estimated using Eq.~\ref{eq:class}.

@@

\textbf {Recurrent Neural Network (RNN)} is a type of Neural Network, which can send feedback from the current hidden state to the hidden state of the next time step. In this way, RNN can capture long-term dependencies for predicting the future events, which is the main advantage of RNN over feedforward Neural Network, MC, and HMM. The capability of remembering past event is very useful in motivational interviews where a behavior observed at some points in the interview is very informative for the future behaviors that will be observed. However, it was observed that RNN fails to capture long-term dependencies due to vanishing gradient problem \cite{bengio1993problem}. In order to mitigate this problem, Hochreiter et al.\cite{hochreiter1997long} proposed a special kind of RNN called Long Short Term Memory networks or simply LSTM. Later, the model was enhanced by including forget gates and the tanh activation function \cite{graves2013speech}. There are several variants of LSTM model, where a dramatic variation on the LSTM is the Gated Recurrent Unit\cite{cho2014properties} or simply GRU. GRUs are simpler than LSTM units and experimentally outperform than other models in many cases. We reiterate the mathematical formulation of GRU defined by Chung et al.\cite{chung2014empirical} as follows (Eq.~\ref{eq:firstgru}~to~\ref{eq:lastgru}):
\begin{equation}
z_t = \sigma(W_zx_t + U_zh_{t-1} + b_z)
\label{eq:firstgru}
\end{equation}
\begin{equation}
r_t = \sigma(W_rx_t + U_rh_{t-1} + b_r)
\end{equation}
\begin{equation}
\tilde h_t = tanh(W_hx_t + r_t \odot U_hh_{t-1} + b_h) 
\end{equation}
\begin{equation}
h_t = z_t \odot h_{t-1} + (1-z_t) \odot \tilde h_t
\label{eq:lastgru}
\end{equation}  
\begin{equation}
total\ loss = \alpha \cdot \frac{1}{T}\sum_{t=1}^T loss(\bar y^{(t)},y^{(t)}) + (1 - \alpha) \cdot loss(\bar y^{(T)},y^{(T)})
\label{eq:loss}
\end{equation}  
\begin{figure}[!htb]
    \centering
    \includegraphics[width=0.70\textwidth]{figures/rnn.eps}
    \caption{\textbf{Proposed RNN model with target replication (TR)}.}
    \label{fig:rnn-model}
\end{figure}
The architecture of the proposed Recurrent Neural Network is shown in Figure~\ref{fig:rnn-model}. We used one hidden layer of 15 LSTM or GRU nodes after embedding the given MYSCOPE codes from the input sequence. Then, we applied a softmax layer at each time-step for predicting the target which is the label of the entire sequence (successful or unsuccessful). This target replication (TR) strategy was used in several studies \cite{lipton2015learning, choi2016doctor} after introduced by Lee et al.\cite{lee2015deeply} in 2015 for convolutional neural networks. As it can be seen from equation~(\ref{eq:loss}), our loss is the combination of final loss and the mean of the losses over all sequence steps, where T is the total number of sequence step, $\bar y$ is the output at step t, and $\alpha\ \epsilon\ [0, 1]$ is a hyperparameter that determines the relative importance of intermediate outputs. Experimentally, it was determined that best performance achieved when we use  $\alpha=0.5$. Our model also contains several other hyperparameters such as embedding dimension, number of hidden units, learning rate, batch size, etc., which were determined by the validation set. We implemented our models in Tensorflow with adam optimizer and early stopping based on the validation loss. Experimentally, we observed that our model converges to optimal after 100 epoch and have no effect of dropout and regularization. The code for all models is publicly available at https://github.com/teanalab/myscope-sequential-analysis.

@@

\section*{Conclusion}
In this paper, we proposed Recurrent Neural Networks with two baseline methods Markov Chain and Hidden Markov Model for predicting the success of motivational interviews. We found out that individual patient-provider communication exchanges are highly indicative of the overall progression and future trajectory of clinical interviews and can be used to predict their overall success. Our proposed method can facilitate motivational interviewing researchers in establishing causal relationships between different communication strategies and the desired behavioral outcomes during the interviews without resource intensive manual qualitative analysis of interview transcripts. Our proposed method can also help to identify most likely sequences that are common to successful and unsuccessful motivational interviews, which can also directly inform clinical practice. This work also has broad implications for qualitative public health research by providing a formal theoretically-grounded computational mechanism to facilitate the development of effective behavioral interventions.

>>>>

In general, there are three categories of classification methods for temporally ordered sequences, e.g.: feature-based, distance metric-based and model-based method. Feature-based classification methods transform a sequence into a feature vector and apply a standard supervised learning algorithm, such as support vector machine \cite{leslie2004fast} or decision tree \cite{chuzhanova1998feature}. Shapelet \cite{ye2009time} and pattern \cite{kudenko1998feature, lesh1999mining} based techniques as well as hierarchical approaches \cite{nallam2016effective} have been proposed instead of standard classifiers as well. Distance-based methods measure the similarity between sequences of behavior codes to determine the quality of classification. The most commonly used distance function is Euclidian distance \cite{keogh2003need} with Dynamic Time Wrapping \cite{keogh2000scaling} used for more flexible matching in time series data. The order of observations in a sequence is challenging to model using feature-based methods because the traditional bag-of-words representation of features does not account for the order of observation. At the same time, distance-based methods are more appropriate for time series data. Model-based classification methods involve building a model based on the dataset. In other words, information was learned from the dataset, and use that as a "model" to classify sequences.

@@

One of the model-based classification methods first creates a probabilistic model of a behavioral code sequence, such as Markov Chain (MC) and Hidden Markov Model \cite{rabiner1989tutorial} (HMM). Markov models have limitations when applied to sequence classification because they cannot preserve long-term dependencies in sequences. Similarly, the hidden states of HMMs follow a first-order Markov chain and can select only one hidden state at each time step. Therefore, Markov chain and HMMs are only used for passing one or two preceding step(s) among sequence steps\cite{kundu1988recognition}. On the other hand, Recurrent Neural Networks (RNNs) have no such limitations in preserving long-term dependencies among sequence steps due to their unlimited memory. It was shown that RNNs are very effective in sequence modeling with temporal data. Excellent performance in a wide range of sequence modeling problems with temporal data has been observed using RNNs \cite{nion2013handwritten, lipton2015learning, choi2016doctor}. A variant of RNN called Long Short-Term Memory was proposed \cite{graves2013speech} which solved the vanishing gradient problem \cite{bengio1993problem}. LSTM showed excellent performance in several domains, including speech recognition \cite{graves2013speech}, handwriting recognition \cite{nion2013handwritten} and health informatics \cite{lipton2015learning, choi2016doctor}. In Lipton et. al.\cite{lipton2015learning}, LSTM was used as a multilabel classification model to recognize patterns in multivariate time series of clinical measurements such as body temperature, heart rate, blood pressure, etc. LSTM was also effectively used, recently, for designing a model, called Doctor AI \cite{choi2016doctor}, to predict the diagnosis and medication categories for a subsequent visit. A further simplification and improvement of the LSTM model, called the Gated Recurrent Unit\cite{chung2014empirical} (GRU), was later proposed. These variants show markedly improved performance among all RNN variants and as such, this work focuses on these variants of RNN specifically.

In this paper, we test the applicability of state of the art models for the classification of patient-provider communication sequences in the context of clinical interviews. Specifically, we focus on the transcripts of Motivational Interviews (MI) with obese adolescents and their caregivers. Childhood obesity is a serious public health concern in the United States and worldwide. Recent estimates indicate that approximately one-third (31.8\%) of US children age 2-19 years are overweight and 16.9\% are obese \cite{ogden2012prevalence}. Adolescents who are obese likely continue to be obese in adulthood and have a greater risk of heart disease, type 2 diabetes, stroke, cancer, and osteoarthritis \cite{general2010surgeon}. One approach to effective obesity intervention is Motivational Interviewing, an evidence-based counseling technique to increase intrinsic motivation and self-efficacy for health-related behavior change. The goal of MI is to encourage patients to explore their own desires, ability, reasons, need for and commitment to the targeted behavior change. These statements, referred to as ``change talk'' (or CT), consistently predict actual behavior change\cite{apodaca2009mechanisms} that can be sustained for as long as 34 months\cite{walker2011influence} after an interview. However, the ability of counselors to consistently elicit this type of patient communication requires knowledge of effective communication strategies for a variety of patients, which can only be obtained through analysis of a large number of annotated interviews. Since manual examination and sequential pattern analysis of transcripts is a very time-consuming process, tailoring of MI interventions to particular populations can take years. Therefore, there is a need for informatics-based methods to facilitate the development of effective MI-based interventions, in general, and theoretically-grounded computational models to explore the mechanisms of MI's efficacy, in particular.

@@

\section*{Methods}
\subsection*{\textit{Data collection}}
The experimental dataset for this work was constructed from the transcripts of 129 motivational interviews, which include a total of 50,239 segmented and annotated utterances. Each transcript consists of an MI interview session involving counselor, adolescent, and caregiver. The utterances were annotated based on MYSCOPE codebook \cite{carcone2013provider}, which are grouped into client codes (adolescent and caregiver) group and counselor codes group. Utterances were divided into successful and unsuccessful communication sequences. Successful communication sequences result in positive change talk and commitment language statements by an adolescent or caregiver, while unsuccessful sequences are the ones that result in negative change talk or commitment language and the sequences, in which no change talk or commitment language statements occur. Out of 5143 observed sequences, 4225 were positive and 918 were negative. Here, the successful sequences had an average length of 9.79 while that of unsuccessful sequences was 9.65. For each of the probabilistic models (Markov chain and HMM), two models were trained, one model was trained using successful sequences and another one was trained using unsuccessful sequences.

@@

Our data set is imbalanced and predominately composed of successful sequences with only a small percentage (17.85\%) of unsuccessful sequences. Usually, predictive accuracy misrepresents the performance of an employed algorithm for an imbalanced data. Because simply predict the majority class (successful) would provide a predictive accuracy 82.15\%. Therefore, it is important to handle imbalance data properly. Many solutions are proposed to deal with the imbalanced data sets, which can be divided into two categories: data level and algorithmic level. In this study, we utilized under and oversampling methods at data processing level for balancing the adolescent obesity data set. Synthetic Minority Over-sampling (SMOTE) method is the most widely used method for over-sampling an imbalanced dataset, in which new synthetic examples were generated from minority class \cite{chawla2002smote}. We generated synthetic examples along the borderline between minority examples and their selected nearest neighbors \cite{nguyen2011borderline}. On the other hand, the undersampling method under samples the majority class by replacing a cluster of majority samples by the cluster centroid of a K-Means algorithm.  A fragment of an adolescent session transcript is presented in Table~\ref{tab:anno_examp}. Hence, annotation column shows the successful sequence of behavior codes from top to bottom, in particular, ($SS\rightarrow OQO\rightarrow HUPO\rightarrow OQTBN\rightarrow CHT+$) is a successful behavior sequence, where counselor starts with a structured session and gets positive change talk at the end. It was noticed that same context ``Yeah" and ``Yes'' represent different behavior codes CHT+ and HUPW, respectively. Because human brain can remember the past behaviors, and classify the current context into different behavior codes after considering their past behaviors. For this reason, we only consider those models that can capture the past events for the classification of behavior sequences.\\

@@

The above model also referred as first-order MC, since it only considers immediately preceding behavior code when computing the state transition probabilities. In our experiment, we also considered second-order Markov model that compute transition probabilities based on previous two states.

\textbf {Hidden Markov Model (HMM)} is another probabilistic model, which is used for modeling processes varying in time. HMMs are widely used for sequence analysis because of their ability to incorporate dependencies among elements in sequence. Mathematically, HMM can be defined as the following: $\lambda = (N, M, A, B, \pi)$
\begin{itemize}
\item N is a number of hidden states in the model
\item M is a number of distinct observations symbols per state, i.e. the discrete vocabulary size
\item A is an $N\times N$ state transition probability distribution matrix $A = \{a_{ij}\}$
\item B is an $N\times M$ matrix $B = \{b_j(k)\}$ with observation symbol probability distribution for each state 
\item $\pi$ is the initial state distribution vector $\pi = \{\pi_i\}$
\end{itemize}
HMM is also represented by a compact notation: $\lambda = (A, B, \pi)$ after excluding the structure parameters M and N. The key difference between HMM and MC is that HMM requires specifying the number of hidden states as a model parameter and then the model deduces a sequence of hidden states that best explains the observations along with state transition probabilities and distributions of observation symbols for each state. The Baum-Welch algorithm was used to estimate the parameters of HMMs for successful and unsuccessful interviews using the corresponding training set, while the Viterbi algorithm was used to determine the most likely sequence of hidden states for a given sequence of observations. After assignment of hidden states, the log-likelihood of successful outcome can be estimated using Eq.~\ref{eq:class}.

@@

\textbf {Recurrent Neural Network (RNN)} is a type of Neural Network, which can send feedback from the current hidden state to the hidden state of the next time step. In this way, RNN can capture long-term dependencies for predicting the future events, which is the main advantage of RNN over feedforward Neural Network, MC, and HMM. The capability of remembering past event is very useful in motivational interviews where a behavior observed at some points in the interview is very informative for the future behaviors that will be observed. However, it was observed that RNN fails to capture long-term dependencies due to vanishing gradient problem \cite{bengio1993problem}. In order to mitigate this problem, Hochreiter et al.\cite{hochreiter1997long} proposed a special kind of RNN called Long Short Term Memory networks or simply LSTM. Later, the model was enhanced by including forget gates and the tanh activation function \cite{graves2013speech}. There are several variants of LSTM model, where a dramatic variation on the LSTM is the Gated Recurrent Unit\cite{cho2014properties} or simply GRU. GRUs are simpler than LSTM units and experimentally outperform than other models in many cases. We reiterate the mathematical formulation of GRU defined by Chung et al.\cite{chung2014empirical} as follows (Eq.~\ref{eq:firstgru}~to~\ref{eq:lastgru}):
\begin{equation}
z_t = \sigma(W_zx_t + U_zh_{t-1} + b_z)
\label{eq:firstgru}
\end{equation}
\begin{equation}
r_t = \sigma(W_rx_t + U_rh_{t-1} + b_r)
\end{equation}
\begin{equation}
\tilde h_t = tanh(W_hx_t + r_t \odot U_hh_{t-1} + b_h) 
\end{equation}
\begin{equation}
h_t = z_t \odot h_{t-1} + (1-z_t) \odot \tilde h_t
\label{eq:lastgru}
\end{equation}  
\begin{equation}
total\ loss = \alpha \cdot \frac{1}{T}\sum_{t=1}^T loss(\bar y^{(t)},y^{(t)}) + (1 - \alpha) \cdot loss(\bar y^{(T)},y^{(T)})
\label{eq:loss}
\end{equation}  
\begin{figure}[!htb]
    \centering
    \includegraphics[width=0.70\textwidth]{figures/rnn.eps}
    \caption{\textbf{Proposed RNN model with target replication (TR)}.}
    \label{fig:rnn-model}
\end{figure}
The architecture of the proposed Recurrent Neural Network is shown in Figure~\ref{fig:rnn-model}. We used one hidden layer of 15 LSTM or GRU nodes after embedding the given MYSCOPE codes from the input sequence. Then, we applied a softmax layer at each time-step for predicting the target which is the label of the entire sequence (successful or unsuccessful). This target replication (TR) strategy was used in several studies \cite{lipton2015learning, choi2016doctor} after introduced by Lee et al.\cite{lee2015deeply} in 2015 for convolutional neural networks. As it can be seen from equation~(\ref{eq:loss}), our loss is the combination of final loss and the mean of the losses over all sequence steps, where T is the total number of sequence step, $\bar y$ is the output at step t, and $\alpha\ \epsilon\ [0, 1]$ is a hyperparameter that determines the relative importance of intermediate outputs. Experimentally, it was determined that best performance achieved when we use  $\alpha=0.5$. Our model also contains several other hyperparameters such as embedding dimension, number of hidden units, learning rate, batch size, etc., which were determined by the validation set. We implemented our models in Tensorflow with Adam optimizer and early stopping based on the validation loss. Experimentally, we observed that our model converges to optimal after 100 epoch and have no effect of dropout and regularization. The code for all models is publicly available at https://github.com/teanalab/myscope-sequential-analysis.

@@

\section*{Conclusion}
In this paper, we proposed Recurrent Neural Networks with two baseline methods Markov Chain and Hidden Markov Model for predicting the success of motivational interviews. We found out that individual patient-provider communication exchanges are highly indicative of the overall progression and future trajectory of clinical interviews and can be used to predict their overall success. Our proposed method can facilitate motivational interviewing researchers in establishing causal relationships between different communication strategies and the desired behavioral outcomes during the interviews without resource-intensive manual qualitative analysis of interview transcripts. Our proposed method can also help to identify most likely sequences that are common to successful and unsuccessful motivational interviews, which can also directly inform clinical practice. This work also has broad implications for qualitative public health research by providing a formal theoretically-grounded computational mechanism to facilitate the development of effective behavioral interventions.
